---
title: "Using R Goslin to parse and normalize lipid nomenclature"
author: "Nils Hoffmann"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This project is a parser, validator and normalizer implementation for shorthand lipid nomenclatures, base on the Grammar of Succinct Lipid Nomenclatures project.

[Goslin](https://github.com/lifs-tools/goslin) defines multiple grammers compatible with ANTLRv4 for different sources of shorthand lipid nomenclature. This allows to generate parsers based on the defined grammars,
which provide immediate feedback whether a processed lipid shorthand notation string is compliant with a particular grammar, or not.

> **_NOTE:_**  This is an *early* development version, please use at your own risk and report issues to help improve it!

rgoslin uses the Goslin grammars and the cppgoslin parser to support the following general tasks:

1. Facilitate the parsing of shorthand lipid names dialects.
2. Provide a structural representation of the shorthand lipid after parsing.
3. Use the structural representation to generate normalized names.

## Related Projects

- [Goslin grammars and reference test files](http://github.com/lifs-tools/goslin)
- [C++ implementation](https://github.com/lifs-tools/cppgoslin)
- [Java implementation](https://github.com/lifs-tools/jgoslin)
- [Python implementation](https://github.com/lifs-tools/pygoslin)
- [Webapplication and REST API](https://github.com/lifs-tools/goslin-webapp)

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
```
### Supported nomenclatures

- LipidMAPS
- SwissLipids
- Shorthand (Liebisch et al.) nomenclature
- Fragments (as full parts)
- HMDB

### Using the library

In order to use the provided translation functions of rgoslin, you first need to load the library.
```{r}
library(rgoslin)
```

To check, whether a given lipid name can be parsed by any of the parsers supplied by cppGoslin, you can use the `isValidLipidName` method. It will return `TRUE` if the given name can be parsed by any of the available parsers and `FALSE` if the name was not parseable.
```{r}
isValidLipidName("PC 32:1")
```

#### Parsing a single lipid name

Using `parseLipidName` with a lipid name returns a named vector of properties of the parsed lipid name.
```{r}
pc32vector <- parseLipidName("PC 32:1")
pc32df <- as.data.frame(t(pc32vector))
```


```{r echo = FALSE, results = 'asis'}
kable(pc32df %>% select(-starts_with("FA"),-starts_with("LCB")), caption = "Lipid name parsing results for PC 32:1, FA and LCB columns omitted for brevity.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```

```{r echo = FALSE, results = 'asis'}
kable(pc32df %>% select(`Normalized Name`, starts_with("FA"),starts_with("LCB")), caption = "Lipid name parsing results for PC 32:1 with FA and LCB columns.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```

If you want to set the grammar to parse against manually, this is also possible:

```{r}
originalName <- "TG(16:1(5E)/18:0/20:2(3Z,6Z))"
tagVec <- rgoslin::parseLipidNameWithGrammar(originalName, "LipidMaps")
tagDf <- as.data.frame(t(tagVec))
```

```{r echo = FALSE, results = 'asis'}
kable(tagDf %>% select(-starts_with("FA"),-starts_with("LCB")), caption = "Lipid name parsing results for TG isomeric subspecies, FA and LCB columns omitted for brevity.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```

```{r echo = FALSE, results = 'asis'}
kable(tagDf %>% select(`Normalized Name`, starts_with("FA"),starts_with("LCB")), caption = "Lipid name parsing results for TG isomeric subspecies with FA and LCB columns.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "200px")
```

#### Parsing multiple lipid names

If you want to parse multiple lipid names, use the `parseLipidNames` method with a vector of lipid names. This returns a data frame of properties of the parsed lipid names with one row per lipid.

```{r}
multipleLipidNames <- parseLipidNames(c("PC 32:1","LPC 34:1","TG(18:1_18:0_16:1)"))

```

```{r echo = FALSE, results = 'asis'}
kable(multipleLipidNames %>% select(-starts_with("FA"),-starts_with("LCB")), caption = "Lipid name parsing results for PC 32:1, LPC 34:1, TG(18:1_18:0_16:1), FA and LCB columns omitted for brevity.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "300px")
```
```{r echo = FALSE, results = 'asis'}
kable(multipleLipidNames %>% select(`Normalized Name`, starts_with("FA"), starts_with("LCB")), caption = "Lipid name parsing results for PC 32:1, LPC 34:1, TG(18:1_18:0_16:1) with FA and LCB columns.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "300px")
```

Finally, if you want to parse multiple lipid names and want to use one particular grammar:

```{r}
originalNames <- c("PC 32:1","LPC 34:1","TAG 18:1_18:0_16:1")
multipleLipidNamesWithGrammar <- parseLipidNamesWithGrammar(originalNames, "Goslin")
```

```{r echo = FALSE, results = 'asis'}
kable(multipleLipidNamesWithGrammar %>% select(-starts_with("FA"),-starts_with("LCB")), caption = "Lipid name parsing results for Goslin grammar and lipids PC 32:1, LPC 34:1, TG(18:1_18:0_16:1), FA and LCB columns omitted for brevity.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "300px")
```

```{r echo = FALSE, results = 'asis'}
kable(multipleLipidNamesWithGrammar %>% select(`Normalized Name`, starts_with("FA"), starts_with("LCB")), caption = "Lipid name parsing results for Goslin grammar and lipids PC 32:1, LPC 34:1, TG(18:1_18:0_16:1) with FA and LCB columns.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "300px")
```

### Parsing Adducts

The Goslin parser also support reading of lipid shorthand names with adducts:

```{r}
originalNames <- c("PC 32:1[M+H]1+","PC 32:1")
lipidNamesWithAdduct <- parseLipidNamesWithGrammar(originalNames, "Goslin")
```

This will populate the columns "Adduct" and "AdductCharge" with the respective values. Please note that we require the adduct and its charge in IUPAC recommended nomenclature: 

```{r echo = FALSE, results = 'asis'}
kable(lipidNamesWithAdduct %>% select(-starts_with("FA"),-starts_with("LCB")), caption = "Lipid name parsing results for Goslin grammar and lipids PC 32:1[M+H]1+ and PC 32:1, FA and LCB columns omitted for brevity.") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", font_size = 7)) %>% scroll_box(width = "800px", height = "300px")
```

### Getting help & support
If you find any issues with the library, would like to have other functionality included, require help or would like to contribute, please contact us via our [GitHub Project](https://github.com/lifs-tools/rgoslin) or via the [LIFS support page](https://lifs.isas.de/support.html).
